Actuator line modeling of vertical-axis turbines

### 2016-08-27 reviews/reviewer2-response.md

>1. Pag1. Please insert the list of Symbols since there are a lot of
abbreviations in the text, a quick reference could be useful.

We have updated the manuscript with a nomenclature section.

>2. Pag. Pag3. Sect. 2. The parameter reduced frequency "k" is the key parameter
for unsteadiness considerations concerning VAT: some more detailed comments
should be given concerning the actual value of this parameter in the case study
numerically modelled. At the same time a comment should be given concerning the
ALM capabilities (and/or future improvements) in terms of unsteady aerodynamics.

We have added values for the reduced frequencies of the rotors tested here, and
a note that the LB dynamic stall models do contain corrections for unsteady
aerodynamics even in the absence of stall. This implies that these models should
work better than simply using static foil data, e.g., at high TSR, where angles
of attack do not exceed the static stall angle.

@@

>4. Pag6. Sect.7. Does the authors took into consideration that the ALM could be
in theory able to catch the Tip vortex effect without a correction model? i.e.
the vortex generated by the lifting of the blade line is already influencing the
angle of attack on the blade.

While tip vortices are produced by the ALM, and these do affect the local angle
of attack, we are not sure how easy it would be to create a ALM that resolves
tip effects totally accurately without any correction model. The issue may be
that since we do not enforce the no peneration BC through the blade, the tip
vortex can leave the blade too early, which causes incorrect angles of attack
close to the tip.

>5. Pag7. Sect.8-9 . Add a sketch of the domain boundaries with the position of
the turbine, and the flow direction. Figure 3 is not very effective in terms of
quantitative information on the actual implementation of the code.

This is a valid point. We have created a new figure to better describe the
computational domain and dimensions.

### 2016-08-27 paper.tex
Add note about reduced frequencies encountered; resolves #32

For a VAT or CFT reduced frequency can be reformulated in terms of the tip speed
ratio as
\begin{equation}
    k = \frac{\lambda c}{2R},
\end{equation}
which is then also a function of solidity or chord-to-radius ratio $c/R$. As an
example, a large scale, relatively low solidity Darrieus wind turbine such as
the Sandia 34 m diameter Test Bed, with an equatorial blade chord of 0.91 m
\cite{Murray2011}, a reduced frequency $k=0.16$ is encountered based solely on
angle of attack oscillations at $\lambda=6$. For a smaller scale CFT, e.g., with
$c/R = 0.25$, operating at $\lambda = 2$, the reduced frequency is 0.25. It
follows that unsteady effects will be significant for cross-flow turbines even
in the absence of nonlinear effects such as stall.

@@

In this study we employed a dynamic stall model developed for low mach numbers
by Sheng et al. \cite{Sheng2008}---derived from the Leishman--Beddoes (LB)
semi-empirical model \cite{Leishman1989}. This model, along with two other
Leishman--Beddoes model variants, was tested for its effectiveness in cross-flow
turbine conditions by Dyachuk et al.~\cite{Dyachuk2014}, who concluded that the
Sheng et al. variant results matched most closely with experiments. In a similar
study \cite{Dyachuk2015}, the Sheng et al. model also performed better than the
Gormont model \cite{Gormont1973}, which inspired its adoption here for the ALM.

>>>>

For a VAT or CFT reduced frequency can be reformulated in terms of the tip speed
ratio as \begin{equation} k = \frac{\lambda c}{2R}, \end{equation} which is then
also a function of solidity or chord-to-radius ratio $c/R$. As an example, a
large scale, relatively low solidity Darrieus wind turbine such as the Sandia 34
m diameter Test Bed, with an equatorial blade chord of 0.91 m \cite{Murray2011},
a reduced frequency $k=0.16$ is encountered based solely on angle of attack
oscillations at $\lambda=6$. For the two turbines studied here, reduced
frequencies at $\lambda_0$ are 0.27 and 0.17 (based on root chord length) for
the UNH-RVAT and RM2, respectively. Therefore, both cases will be inherently
unsteady, with these effects dominating the UNH-RVAT's behavior. Hence, an
accurate unsteady aerodynamics model is important to predicting blade loading.

@@

In this study we employed a dynamic stall model developed for low mach numbers
by Sheng et al. \cite{Sheng2008}---derived from the Leishman--Beddoes (LB)
semi-empirical model \cite{Leishman1989}. This model, along with two other
Leishman--Beddoes model variants, was tested for its effectiveness in cross-flow
turbine conditions by Dyachuk et al.~\cite{Dyachuk2014}, who concluded that the
Sheng et al. variant results matched most closely with experiments. In a similar
study \cite{Dyachuk2015}, the Sheng et al. model also performed better than the
Gormont model \cite{Gormont1973}, which inspired its adoption here for the ALM.
Note that even in the absence of stall, the LB DS models still modulate foil
force and moment coefficients to account for unsteadiness, which is important
for the reduced frequencies of the rotors simulated here.
